# Abstract

```{r options_allo, include = FALSE}
opts_chunk $ set (fig.path = 'Figures/', dev = 'pdf')
```

```{r captions_allo, include = FALSE}

captions $ fig.cac_logCS <- "**Allometric shape variation.** (a) Shape variation associated with the Common Allometric Component (CAC). (b) Random regression of projections on the CAC over log Centroid Size. A regression line is adjusted for each species, and dots represent their mean values. \\label{fig:cac_logCS}"

captions $ fig.phylo_W <- "**Deviations from ancestral allometry.** (a) Intercepts; (b) Slopes. Asterisks indicate whether a given deviation is distinct from zero, considering 95\\% credible intervals. \\label{fig:phylo_W}"

captions $ fig.MI_vs_slopeW_main <- "**Allometry and modularity.** Modularity Hypothesis Index (MHI) regressed over allometric slope deviations ($b_s$). For each global trait set (Face and Neurocranium), the shaded region around regression lines indicate 95\\% credible intervals around estimated regression parameters. \\label{fig:MI_vs_slopeW_main}"

if (thesis) {
 
  captions $ fig.MI_vs_slopeW_si <- "Modularity Indexes (MI) *vs.* allometric slope deviations ($b_s$). For each local trait set, the shaded region around regression lines indicate 95\\% credible intervals around estimated regression parameters. \\label{fig:MI_vs_slopeW_si}"

}

```

# Introduction

Anthropoid primates display an approximately thousand-fold variation of body sizes, from pygmy marmosets ($\sim 110g$) to lowland gorillas ($\sim 120kg$).
This astonishing size diversity implies substantial morphological, physiological and ecological differences among these organisms, since changes in body size fundamentally affect how living organisms (and their composing parts) interact with their environments [@gould_allometry_1966; @west_general_1997].
These differential scaling relationships between organismal traits and body size are referred to as allometry [@huxley_problems_1932; @gould_allometry_1966; @cheverud_relationships_1982; @lande_genetic_1985] and may be represented as power laws, which become linear on a log-log scale [@huxley_problems_1932; but see @nijhout_developmental_2012].
Allometry may thus be represented by two parameters: (allometric) slopes and intercepts.

Allometry may act as a constraint, diverting evolutionary trajectories in natural populations [@gould_spandrels_1979; @lande_quantitative_1979; @schluter_adaptive_1996].
For example, selection for body size driven by shifts in feeding habits has dominated diversification in New World monkeys, and differences in skull morphology among species in this group appear to be merely an indirect consequence of the allometric relationships between skull traits and body size [@marroig_size_2005; @marroig_size_2010; @marroig_modularity_2012].
With respect to Old World Monkeys, skull morphological evolution may also be adaptive and related to allometric associations, at least within Cercopithecids [@cardini_geometric_2007; @cardini_variation_2008]; regarding Hominids, selection for reduction of facial elements and increased neurocranial capacity has played a limited role in the divergence between *Homo* and *Australopithecus* [@ackermann_detecting_2004; @schroeder_characterizing_2014]

Regarding morphological traits, allometric slopes may be constrained throughout development due to stabilizing selection [@lande_genetic_1980; @cheverud_quantitative_1984; @cheverud_developmental_1996; @hansen_epigenetics:_2011; @zelditch_epigenetic_2011], while intercepts may be more labile on an evolutionary scale [@gould_allometry_1974; @egset_artificial_2012; @bolstad_complex_2015].
However, these parameters are dependant upon the mean individual growth rate and the covariance between these growth rates and body size achieved at adulthood; thus, selection on body size may also change allometric slopes [@pelabon_relationship_2013; @voje_allometric_2013].
Experimental evidence on the evolution of allometric relationships has accumulated, indicating that growth parameters harbor genetic variation and thus may be available to selection, an effect that will then cascade to populational allometric relationships [e.g. @frankino_natural_2005; @tobler_developmental_2010; @egset_artificial_2012], although methodological issues may hamper some of this evidence [@houle_measurement_2011; @voje_allometric_2013].
Therefore, whether and to which extent allometric slopes can evolve remain as open questions.

In the present work, we investigate the relationship between skull size and shape in a phylogenetic framework using Anthropoid Primates as model organisms.
We focus on testing whether and to which proportion intercepts and slopes have changed during the diversification of this group.
Moreover, integration patterns in skull form are quite stable in both New World [@marroig_comparison_2001] and Old World Monkeys, with some exceptions [@oliveira_covariance_2009].
Several authors have argued that allometric constraints contribute to integration patterns [e.g. @porto_evolution_2009; @armbruster_integrated_2014], but this association has never been formalized.
Thus, we test the hypothesis that variation in allometric parameters, if present, will contribute to variation in the strength of association between skull traits.
Considering both the spatial and temporal dynamics of mammalian skull development [@zelditch_ontogenetic_1989; @hallgrimsson_deciphering_2009; @cardini_larger_2013], we expect that such changes will affect Facial and Neurocranial traits in opposing ways, due to the distinction between early- and late-developmental factors in skull development.

Using a Bayesian phylogenetic random regression model of allometric shape against size, we estimate allometric parameters along the phylogeny of Anthropoidea.
We show that allometric slopes have remained quite stable throughout Anthropoid diversification, while allometric intercepts have suffered changes in several groups, and that the little variation associated with allometric slopes produces opposite changes in Facial and Neurocranial integration, providing a clear link between allometry and integration.
Despite the general stability in allometric slopes, we find remarkable changes occurring in *Homo* and *Gorilla*, demonstrating that constraints imposed by allometry are in no way absolute and, in some situations, may be overcome.

```{r methods_thesis, child = 'allo_meth.Rmd', eval = thesis}
```

# Results

We used the Common Allometric Component [CAC; @mitteroecker_comparison_2004], which represents a pooled estimate of skull shape allometry [represented as local shape variables; @marquez_measurement_2012] as a proxy for the ancestral allometric relationships for Anthropoid Primates. This axis is associated with positive loadings for Facial traits and negative loadings for Neurocranial traits, thus representing a contrast between these regional sets (\autoref{fig:cac_logCS}a). 

```{r cac_logCS, echo = FALSE, fig.cap = captions $ fig.cac_logCS, fig.width = 12, fig.height = 9.8}
allo.Plots $ CAC.logCS
```

To investigate whether species deviate from such ancestral relationships, we carried out a Bayesian mixed model, considering that slopes and intercepts for the regression between individuals projected over the CAC and log Centroid Size (logCS) within each species (\autoref{fig:cac_logCS}b) are codependent upon their phylogenetic relationships. The posterior distribution of deviations from mean for both parameters estimated in this manner (\autoref{fig:phylo_W}) indicates that intercepts deviate from the mean in at least seven lineages: increases occur in *Pithecia* and *Callicebus* within Pitheciids, in the clade composed of Callithrichinae and Aotinae, and in *Alouatta* within Atelids, while lower intercepts are found in the clade composed of *Homo* and *Pan* within Hominidae, in *Hylobates* within Hylobatids, and also in *Ateles* within Atelids (\autoref{fig:phylo_W}a). Allometric slopes for terminals, on the other hand, deviate from the mean in only three species: *Homo sapiens* and both representatives of *Gorilla*, all with slopes shallower than the mean (\autoref{fig:phylo_W}b).

```{r phylo_new, echo = FALSE, fig.height = 12, fig.width = 10, fig.cap = captions $ fig.phylo_W, message = FALSE}
Tree.Plots $ interslope
```

The association between allometric parameters and modularity was measured using regression models of Modularity Indexes for two regions (Face and Neurocranium) and three sub-regions within each region (Oral, Nasal and Zygomatic in the Face; Orbit, Vault and Basicranium in the Neurocranium) on the estimated allometric parameters. Model selection using the Deviance Information Criterion [DIC; @gelman_bayesian_2004] show that models considering only the intercept always have the worst fit. In most cases, models including only the slope are better than or equal ($\Delta DIC < 2$) to models with both parameters (intercept and slope), suggesting a negligible effect of the intercept. For the Neurocranium region and the Vault sub-region, the model including intercept and slope shows a better fit than the slope-only model; in both cases, however, the effects of the intercept are not significant when one takes into account the posterior distribution of the regression coefficients (results not shown). We thus consider that the models including only allometric slopes provide the best representation of the association between allometric parameters and modularity indexes (\autoref{tab:dic_allo_im}). It should be noted that none of the models tested for the Orbit and Basicranium sub-regions show any association between allometric parameters and Modularity Indexes, therefore these sub-regions were not included in the model selection. Allometric slopes show opposite effects on the Modularity Index for Face and Neurocranium (\autoref{fig:MI_vs_slopeW_main}), and,  with the exception of the Orbit and Basicranium, sub-regions follow the same pattern as the more inclusive partition in which they are contained (\autoref{fig:MI_vs_slopeW_si}).

`r if (thesis) "\\input{Tables/dic_allo_im.tex}"`

```{r MI_vs_slopeW_main, echo = FALSE, fig.width = 7, fig.height = 4, fig.cap = captions $ fig.MI_vs_slopeW_main}
allo.Plots $ MI.vs.Wslope.main
```

```{r MI_vs_slopeW_si, echo = FALSE, fig.width = 7, fig.height = 8, fig.cap = captions $ fig.MI_vs_slopeW_si, eval = thesis}
allo.Plots $ MI.vs.Wslope.si
```

# Discussion

The hypothesis that allometry may act as a constraint, deflecting changes in morphological, physiological and other organismal aspects along a size-trait gradient [e.g. @gould_allometry_1974] has historically been confronted with the opposing view that allometry may be in fact much more dynamic and evolvable [e.g. @kodric-brown_allometry_2006; @bonduriansky_sexual_2007; @frankino_experimental_2009]. As recently stated by some authors [@voje_allometric_2013; @pelabon_evolution_2014], this seeming paradox may in large part owe its existence to different meanings for allometry: one related to any monotonic relationship between variables, while the other specifically deals with power-law relationships between variables as defined by @huxley_problems_1932. This imprecision of definition greatly hampers evidence in favor of a dynamic interpretation of allometry; considering the few studies estimating evolvability of narrow-sense allometry, only one clearly demonstrates persistent changes in allometric slopes [@tobler_developmental_2010].

The results we obtained from the phylogenetic random regression model indicate that substantial changes have occurred on intercepts, especially in New World Monkeys (\autoref{fig:phylo_W}a). As we expected, the role of natural selection on the evolution of body size, as explored elsewhere [@marroig_size_2005; @marroig_size_2010; @marroig_modularity_2012] can be related to these changes. For example, in the lineage composed of Callithrichines and *Aotus*, selection for reduced body size is pervasive, and it has been associated with shifts in pre-natal growth rates, considering a maintenance in relative gestation lengths with respect to their sizes [@marroig_size_2009]; such change in the timing of developmental events may be responsible for the changes in intercepts we observe in tamarins, marmosets and owl monkeys. For Atelids, while selection for increased body size has played a role in the differentiation between this lineage and its sister group Cebidae, the shifts in allometric intercepts occurred within this clade in two different lineages, and in opposing directions. 

It is also noteworthy that, in some lineages of New World Monkeys, selection for changes in body size did not affect intercepts or slopes at all; with respect to *Cebus*, for instance, selection for increased body size [@marroig_modularity_2012] produced only a correlated response on skull shape, along the same allometric relationships observed in its sister genus, *Saimiri* [@marroig_when_2007]. The converse is also true: while saki monkeys (*Pithecia*) show increased intercepts, this group has diversified mostly as a consequence of drift, at least with respect to skull morphology [@marroig_cranial_2004].

We have detected significant divergence in allometric slopes in only two lineages, *Homo* and *Gorilla* (\autoref{fig:phylo_W}b), although in other lineages, such as *Alouatta* and *Callithrix*, some changes may also have occurred. Even though *Homo* and *Gorilla* are closely related and exhibit slopes that are shallower than remaining Anthropoids, such slopes represent two distinct patterns of ontogenetic trajectories.
While for *Gorilla* this change mostly represents an increase in neurocranial capacity close the end of such trajectory, for *Homo* it represents deviations occurring since the beginning of post-natal growth in both Facial and Neurocranial traits [@mitteroecker_evolutionary_2008], which results in an dissociation between cranial shape and size in this lineage, when compared to other Hominoids.

Considering that the early evolution of *Homo* has involved directional selection for both increased neurocranial capacity and reduced facial elements at least in its initial stages of diversification [@ackermann_detecting_2004; @schroeder_characterizing_2014] and given the ancestral allometric relationships for Anthropoids (\autoref{fig:cac_logCS}a), this pattern of selection would imply in a correlated response for reduction in cranial size [@lande_quantitative_1979].
Thus, since cranial size for *Homo* is comparable to its sister lineages (\autoref{fig:cac_logCS}b), this shift in allometric relationships may be indicative of the joint effect of directional selection for these cranial shape features and stabilizing selection for cranial size, given that these changes in correlation patterns between size and shape we observe in *Homo* may be the result of shifts in the relationship between both features in the adaptive landscape for this lineage [@arnold_understanding_2008; @jones_mutation_2007; @jones_effects_2012; @melo_directional_2015].

Concurrently, these changes in *Homo* may also be the result of changes in the environmental effects for both Facial and Neurocranial traits, since part of the integration observed in both regions can be attributed to the effect of post-natal muscular activity over bone growth, induced by masticatory loadings [@hallgrimsson_mouse_2008; @zelditch_building_2009; @herring_muscle-bone_2011].
Given that the act of processing and cooking meals may already have produced changes in physiological aspects of food intake in humans [@wrangham_cooking_2003], this cultural aspect of our ecology may also imply in shallower allometric relationships, since food processing and cooking reduces masticatory loadings associated with its intake. 

Changes in allometric slopes throughout Anthropoid diversification imply in changes in the strength of association between Facial and Neurocranial traits, in opposing ways (\autoref{fig:MI_vs_slopeW_main}).
This indicates that, although the distinction between these two regions is the by-product of developmental interactions since the initial steps of skull ontogeny [@hallgrimsson_mouse_2008], allometric relationships reinforce the effect of such interactions due to the induction of facial growth by muscular activity from weaning to adulthood [@zelditch_ontogeny_1992; @zelditch_building_2009].
Thus, groups with steeper slopes, such as *Ateles* and *Alouatta*, exhibit strong Facial integration and lower Neurocranial integration, while the converse is true for groups with shallower slopes, such as Callthrichines, *Homo*, and *Gorilla*.
Although the bulk of this effect over phenotypic variation may have an environmental origin [@cheverud_relationships_1982], muscular activity is necessary for the proper development of osteological elements [@herring_muscle-bone_2011], thus indicating that even though such functional interactions are dependant on the environment, they are canalized through skull ontogeny in a predictable manner, since developmental systems themselves are under selection for robustness and replicability [@wagner_perspective:_1996; @hansen_epigenetics:_2011], exhibiting behaviors similar to machine learning algorithms [@watson_evolution_2013].

Here, we demonstrated that allometric slopes are more constrained to evolve than intercepts in a macroevolutionary scale, under a comparative framework.
The approach we use here to estimate allometric parameters benefits from properly separating variation within and among populations, considering the interdependence arising from phylogenetic relatedness simultaneously; for instance, the same approach could be used to evaluate variation in allometric parameters between individuals in a population whose genealogy is available, enabling us to quantify the influence of both genetic and environmental factors over these parameters.
In fact, given the availability of a strong theoretical background relating allometries defined at different levels of biological organization [@pelabon_relationship_2013], further work in understanding the genetic basis of variation in allometry and how developmental systems interact with such variation may be critical to fully characterize in which conditions such feature evolves.

```{r methods_paper, child = 'allo_meth.Rmd', eval = !thesis}
```
