---
title: "Modularity and Morphometrics: Error Rates in Hypothesis Testing" 
author: Guilherme Garcia^1,2^, Felipe Bandoni de Oliveira^1^ & Gabriel Marroig^1^
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document:
    includes:
      in_header: Tex/modcomp_head.tex
      before_body: Tex/chap_info.tex
    fig_caption: true
    keep_tex: true
  word_document:
    fig_caption: true
    fig_width: 6
fontsize: 12pt
csl: Bib/systematic-biology.csl
bibliography: Bib/modcomp.bib
---	

^1^Laboratório de Evolução de Mamíferos, Departamento de Genética e Biologia Evolutiva, Instituto de Biociências, Universidade de São Paulo, CP 11.461, CEP 05422-970, São Paulo, Brasil

^2^<wgar@usp.br>

running title: Modularity, Error and Morphometrics

key words: $P$-matrix; modularity; primates; development;

```{r, echo = FALSE}
thesis <- FALSE
```

```{r modcomp_body, child = 'modcomp_body.Rmd'}
```

### Acknowledgements

We thank A. Porto, D. A. R. Melo, and G. Burin for comments on an early draft. This work has been funded by CNPq (Conselho Nacional de Pesquisa e Desenvolvimento Tecnológico) and FAPESP (Fundação de Apoio à Pesquisa do Estado de São Paulo).

# References

