# Abstract

The study of modularity in morphological systems has increased in the past twenty years, parallel to the popularization of geometric morphometrics, which has led
to the emergence of different criteria for detecting modularity on landmark data.
However, compared to usual covariance matrix estimators, Procrustes estimators have properties that hinder their use.
Here, we compare different representations of form, focusing on detecting modularity patterns defined *a priori*; we also compare two metrics: one derived from traditional morphometrics (MHI) and another that emerged in the context of landmark data (RV).
Using Anthropoid skulls, we compare these metrics over three representations of form: interlandmark distances, Procrustes residuals, and local shape variables.
Over Procrustes residuals, both tests fail to detect modularity patterns, while in remaining representations they show the distinction between early and late development in skull ontogeny.
To estimate type I and II error rates, we built covariance matrices of known structure; these tests indicate that, considering both effect and sample sizes, tests using MHI are more robust than those using RV.
However, both metrics have low statistical power when used on Procrustes residuals; thus, we conclude that the influence of development and function is poorly represented on Procrustes estimators for covariance matrices.

```{r options_modcomp, include = FALSE}
opts_chunk $ set (fig.path = 'Figures/', dev = 'pdf', fig.pos = 'tbp')
```

```{r captions_modcomp, include = FALSE}

captions $ fig.cor_dist <- 'Distribution of within-set and between-set mean correlations derived from the six types of empirical correlation matrices. \\label{fig:cor_dist}'

captions $ fig.stat_dist_sim <-
  'Distribution of Modularity Hypothesis Index and RV Coefficient for theoretical correlation matrices. \\label{fig:stat_dist_sim}'

captions $ fig.Func <-
  'MHI (a) and RV (b) values for localized trait sets. Circles indicate whether a trait set is recognized as a variational module in a given OTU, with $P(\\alpha)$ indicated by the legend. Notice that values represented in blue are either high MHI values and low RV values, as the alternate hypothesis for each statistic is formulated accordingly; see \'Methods\' for details. \\label{fig:Func}'

captions $ fig.Dev <-
  'MHI (a) and RV (b) values for regional trait sets. Circles indicate whether a trait set is recognized in a given OTU, with $P(\\alpha)$ indicated by the legend. Notice that values represented in blue are either high MHI values and low RV values, as the alternate hypothesis for each statistic is formulated accordingly; see \'Methods\' for details. \\label{fig:Dev}'

captions $ fig.type1 <-
  'Type I error rates as a function of the chosen significance level regarding tests for variational modularity applied on $\\mathbf{C}_r$ correlation matrices. The solid black line represents the identity relationship. \\label{fig:type1}'

captions $ fig.type2_ed <-
  'Power for both MHI and RV statistics as a function of the chosen significance levels with respect to tests for variational modularity applied on $\\mathbf{C}_s$ correlation matrices sampled from the distribution of correlations between interlandmark distances. Lines are colored with respect to quantiles of the $b^2$ distribution. The solid black line represents the identity relationship. \\label{fig:type2_ed}'

captions $ fig.type2_def <-
    'Power for both MHI and RV statistics as a function of the chosen significance levels with respect to tests for variational modularity applied on $\\mathbf{C}_s$ matrices sampled from the distribution of correlations between local shape variables. Lines are colored with respect to quantiles of the $b^2$ distribution. \\label{fig:type2_def}'

captions $ fig.type2_sym <-
  'Power for both MHI and RV statistics as a function of the chosen significance levels with respect to tests for variational modularity applied on $\\mathbf{C}_s$ matrices sampled from the distribution of correlations between Procrustes residuals. Lines are colored with respect to quantiles of the $b^2$ distribution. The solid black line represents the identity relationship. \\label{fig:type2_sym}'

```

# Introduction

Modularity is a characteristic property that biological systems exhibit regarding the distribution of interactions between their composing elements; that is, in a given system, subsets of elements, denominated modules, interact more among themselves than with other such subsets [@newman_modularity_2006; @mitteroecker_conceptual_2007; @wagner_road_2007].
This property has been well documented at different levels of biological organization, from the dynamics of metabolic networks [e.g. @ravasz_hierarchical_2002; @andrade_detecting_2011] to the structure of interactions among individuals in populations [e.g. @fortuna_spatial_2008] and among species in ecological communities [e.g. @genini_cheaters_2010].

Regarding morphological systems, the concept of modularity is associated with the framework of morphological integration [@olson_morphological_1958; @cheverud_developmental_1996], which refers to the organization of covariances or correlations between morphological elements and the hypotheses concerning their relationships.
In this context, modularity is related to the uneven distribution of genetic effects over phenotypic variation articulated through development [genotype/phenotype map; @wagner_homologues_1996]; in a classical quantitative genetics view, these genetic effects are the result of pleiotropy and linkage disequilibrium [@falconer_introduction_1996; @lynch_genetics_1998]. 
A genotype/phenotype map composed of clusters of genes that affect clusters of traits (with little overlap) exhibits a modular organization; this structure is thought to emerge as the result of selection for distinct demands [@wagner_perspective:_1996; @espinosa-soto_specialization_2010; @rueffler_evolution_2012; @melo_directional_2015].
For instance, the decoupling between fore- and hindlimb function in certain mammalian lineages such as bats [@young_serial_2005] and apes [@young_development_2010] is associated with the modularization of both structures, as shown by reduced phenotypic correlations between fore- and hindlimbs and increased correlations between elements within these limbs.

The recognition of variational modules [@wagner_perspective:_1996; @wagner_road_2007] using covariance or correlation patterns in adult populations involves an uderstanding of the underlying developmental and functional dynamics among morphological traits [@polly_developmental_2008; @zelditch_epigenetic_2011].
Skull development in mammals occurs through a series of steps, such as neurocranial growth induced by brain development, and growth mediated by muscle-bone interactions, with spatiotemporal overlapping between such steps [@hallgrimsson_mouse_2008; @herring_muscle-bone_2011; @cardini_larger_2013].
Both timing and scope of each step is regulated by different profiles of genetic expression exhibited by cells originated from different embryonic precursors and their response to signaling factors expressed at the regional level; the response to these signaling factors further changes cell expression profiles, thus generating a feedback loop of diferentiation [@turing_chemical_1952; @marcucio_molecular_2005; @meinhardt_models_2008; @hallgrimsson_deciphering_2009; @franz-odendaal_epigenetics_2011; @minelli_principle_2011].
Each step in this temporal hierarchy may be regarded as modular, since they affect a coherent subset of tissues more so than others [@hallgrimsson_mouse_2008], although each step affect adjoining regions through interactions among developing tissues [@cheverud_effects_1992; @lieberman_epigenetic_2011; @esteve-altava_beyond_2014]; thus, the overlapping of such processes throughout development may complicate their association with correlation patterns [@hallgrimsson_deciphering_2009].

Furthermore, variation in growth rates, which emerges into size variation [@pelabon_relationship_2013; @porto_size_2013], has a particular importance in the context of mammalian morphological systems.
Here size variation refers to variation in both scale (isometric variation) and scale relationships (allometric variation).
This source of variation affects the overall level of correlations between morphological traits [@wagner_eigenvalue_1984; @young_serial_2005], and the magnitude of integration has important consequences for both the evolution of mean phenotypes  [@schluter_adaptive_1996; @marroig_size_2005; @marroig_size_2010; @cardini_larger_2013] and the evolution of morphological integration itself [@oliveira_covariance_2009; @porto_evolution_2009; @shirai_skull_2010; @porto_size_2013].

In this context, adaptive landscapes may be the central component governing both the stability and divergence in integration patterns, as both stabilizing and directional  selection have been shown to produce changes in integration patterns [@arnold_understanding_2008; @jones_mutation_2007; @jones_effects_2012; @melo_directional_2015].
The empirical evidence available demonstrates that both stability [e.g. @marroig_comparison_2001; @oliveira_covariance_2009; @porto_evolution_2009; @willmore_comparison_2009] and divergence [e.g. @monteiro_adaptive_2010; @grabowski_divergent_2011; @sanger_roles_2012; @haber_evolution_2015] of integration patterns are possible outcomes of the evolutionary process.
Therefore, the question of whether any of these two scenarios is the rule or exception at macroevolutionary scales remains open, although some theoretical and methodological differences between these works with respect to the representation of morphological features need to be considered.

### Morphometrics

Traditionally, morphological features have been measured using distances among elements defined in general terms, such as "cranial length" or "cranial width".
@pearson_biometric_1924 introduced the notion that measurements should be restricted to single osteological elements, preferably as distances between homologous features that could be identified in a wide taxonomic coverage.
@cheverud_phenotypic_1982 accomodates this notion into the framework of morphological integration, thus considering individual measurements over single bones as local representations of regional phenomena, that is, the functional, developmental and genetic interactions that produce covariances among these elements.
The influence of such interactions over covariance or correlation matrices can be accessed by defining a subset of measurements which are associated with particular processes and estimating a metric that summarizes such partitioning; the null hypothesis that this partition is undistinguishable from randomly-defined partitions can then be tested using Monte Carlo methods [@mantel_detection_1967; @cheverud_methods_1989]. 

In the past three decades, geometric morphometrics [@bookstein_foundations_1982; @bookstein_morphometric_1991; @goodall_procrustes_1991; @kendall_shape_1984; @rohlf_extensions_1990] has been consolidated as a quantitative framework for the representation of biological shape as geometric configurations of homologous features [landmarks; @bookstein_morphometric_1991].
Two principles are central here: first, the conceptual and statistical separation of size and shape as components of biological form; second, the use of superimposition-based methods [GPA: Generalized Procrustes Analysis; @rohlf_extensions_1990; @goodall_procrustes_1991] for the estimation of shape statistical parameters, such as mean shape and shape covariance structure.
Procrustes estimators were proposed as a solution to the problem that landmark configurations are arbitrarily rotated and translated; such nuisance parameters are impossible to estimate without any assumptions [@goodall_procrustes_1991; @lele_invariance_2002], a situation known as the identifiability problem [@neyman_consistent_1948]. 

Although the use of Procrustes estimators is currently widespread in geometric morphometrics toolboxes [e.g. @klingenberg_morphoj:_2011; @adams_geomorph:_2013], critics have been made regarding the estimation of both mean shape configurations [e.g. @lele_euclidean_1993; @kent_consistency_1997; @huckemann_meaning_2012] and shape covariance matrices [@walker_ability_2000; @adams_geometric_2004; @linde_inferring_2009; @marquez_measurement_2012].
For configurations in two-dimensional space, Procrustes estimators for mean shape perform well under isotropic landmark covariance structure [@kent_consistency_1997], a situation of null covariances among landmarks and coordinates; however, when this assumption does not hold, Procrustes estimates for mean shape may lose either precision and accuracy, especially when shape variation is high [@huckemann_inference_2011].
The example provided by @linde_inferring_2009 demonstrates that when such assumption is broken shape covariance patterns are also poorly estimated; if the unknown landmark covariance matrix is structured due to regional differences in covariance-generating processes, such variation might be displaced and effectively spread out through the entire landmark configuration.

A number of alternatives for estimating shape covariance matrices have already been suggested; some of these alternatives [e.g. @monteiro_evolutionary_2005; @theobald_empirical_2006; @linde_inferring_2009; @zelditch_building_2009] propose modifications to the Procrustes analysis in order to deal with heterogeneity in landmark covariance structure. 
@marquez_measurement_2012 propose another definition of shape descriptors using interpolation-based techniques [@cheverud_finite-element_1986; @bookstein_principal_1989] as a starting point. These descriptors refer to infinitesimal expansions or retractions in reference to a unknown mean shape [@woods_characterizing_2003] estimated at definite locations amidst sampled landmarks.
The authors argue that such descriptors are proper local measurements of shape variation, as they can be directly linked to biological processes that generate covariation among morphological elements.

Despite these caveats, the use of Procrustes estimators for investigating aspects of morphological integration has increased in the past ten years [e.g. @klingenberg_integration_2004; @drake_large_2010; @goswami_influence_2010; @martinez-abadias_pervasive_2011; @sanger_roles_2012].
The results found by these authors are sometimes in stark constrast with similar works using interlandmark distances [e.g. @cheverud_pleiotropic_1997; @oliveira_covariance_2009; @porto_evolution_2009].
For instance, @martinez-abadias_pervasive_2011 found a pattern of strong integration among partitions in human skull covariance patterns, while @oliveira_covariance_2009 and @porto_evolution_2009 demonstrated that humans are among the most modular examples of mammalian skull covariance patterns.
Likewise, while @cheverud_pleiotropic_1997 had shown that 70\% of the pleiotropic effects are confined to either anterior or posterior mandibular components, @klingenberg_integration_2004 has found no evidence for a modular distribution of pleiotropic effects among the partitions of the mouse mandible using the same strain of intercrossed mice at the same generation.
Furthermore, aside from the issues regarding Procrustes estimators, these works also propose different methods to quantify the effects of interactions over covariance patterns. For example, @martinez-abadias_pervasive_2011 and @sanger_roles_2012 use the RV coefficient, a multivariate correlation coefficient defined by @escoufier_traitement_1973 which has been used to quantify modular relationships over landmark covariance patterns since @klingenberg_morphometric_2009 has proposed its use in this context.

### Objectives

In the present work, we compare the methods described by @cheverud_methods_1989 and @klingenberg_morphometric_2009 to test *a priori* defined modularity patterns using Anthropoid Primates as a model organism.
In order to compare the performance of these methods with respect to different representations of form, individuals in our sample are represented both as interlandmark distances and shape variables.
Furthermore, we used an approach based on the construction of theoretical covariance matrices; such matrices are used in order to estimate Type I and Type II error rates for both methods.
This work aims to analyze both methods, which have been proposed in different contexts, under the same conceptual and statistical framework, in order to produce meaningful comparisons.

# Methods

## Sample

The database we used here (\autoref{tab:modcomp_otu}) consists of 21 species, distributed across all major Anthropoid clades above the genus level.
We selected these operational taxonomic units (OTU) from a broader database [@marroig_comparison_2001; @oliveira_covariance_2009] in order to reduce the effects of low sample sizes over estimates of modularity patterns.
Individuals in our sample are represented by 36 registered landmarks, measured using a Polhemus 3Draw (for Platyrrhini) and a Microscribe 3DS (for Catarrhini). Twenty-two unique landmarks represent each individual (\autoref{fig:landmarks}, \autoref{tab:lms}), since 14 of the 36 registered landmarks are bilaterally symmetrical. For more details on landmark registration, see @marroig_comparison_2001 and @oliveira_covariance_2009.

\input{Tables/modcomp_otu}

For each OTU, we estimated phenotypic covariance and correlation matrices for three different types of variables: tangent space residuals, estimated from a Procrustes superimposition for the entire sample, using the set of landmarks described on both \autoref{tab:lms} and \autoref{fig:landmarks} (henceforth Procrustes residuals); interlandmark distances, described in \autoref{tab:dist}; and local shape variables [@marquez_measurement_2012], which are measurements of infinitesimal log volume transformations between each sample unit and a reference (mean) shape, based upon an interpolation function that describes shape variation between sampled landmarks.
In this context, we used thin plate splines as interpolating functions [@bookstein_principal_1989].
We obtained 38 transformations corresponding to the locations of the midpoints between pairs of landmarks used to define interlandmark distances, in order to produce a dataset that represents shape [i.e., form without isometric variation; @bookstein_morphometric_1991; @zelditch_geometric_2004] while retaining the overall properties of the interlandmark distance dataset, such as dimensionality for example.
Furthermore, we were able to use the same hypotheses of trait associations for both types of variables since the position of local shape variables through the skull mirrors the position of interlandmark distances, although they are conceptually different types of measurements.

Here we considered only covariance or correlation structure for the symmetrical component of variation; therefore, prior to any analysis, we controlled the effects of variation in assymmetry. 
For interlandmark distances, we averaged bilateral measurements within each individual.
For both Procrustes residuals and local shape variables, we followed the procedure outlined in @klingenberg_shape_2002 for bilateral structures by obtaning for each individual a symmetrical landmark configuration, averaging each actual shape with its reflection along the sagittal plane; we estimate local shape variables afterwards.
With respect to Procrustes residuals, landmarks placed along the sagittal plane will have zero variation in the direction normal to this plane; we aligned all specimens' sagittal plane to the $xz$ plane, thus removing the $y$ component for each of these landmarks from covariance/correlation matrices.

For each dataset, we estimated covariance and correlation matrices after removing effects of little interest in the present context, such as sexual dimorphism, for example.
For interlandmark distances and local shape variables these effects were removed through a multivariate linear model adjusted for each species, according to \autoref{tab:modcomp_otu}; for Procrustes residuals, the same effects were removed by centering all group means to each species' mean shape since the loss of degrees of freedom imposed by the GPA prohibits the use of a full multivariate linear model over this kind of data to remove fixed effects.

In order to consider the effects of size variation on modularity patterns we used  different procedures to remove the influence of size from each type of variable.
For interlandmark distances we used the approach established by @bookstein_morphometrics_1985; if $\mathbf{C}$ is a correlation matrix, we obtained a correlation matrix $\mathbf{R}$ without the effect of size using the equation
\begin{equation}
\mathbf{R} = \mathbf{C} - \lambda_1 v_1 v^{t}_1
\label{eq:iso}
\end{equation}
where $\lambda_1$ and $v_1$ refer respectively to the first eigenvalue and eigenvector of the spectral decomposition of $\mathbf{C}$, since this eigenvector commonly represents size variation in mammals, especially when interlandmark distances are considered [@wagner_eigenvalue_1984; @mitteroecker_comparison_2004; @mitteroecker_conceptual_2007]; $t$ denotes matrix transpose.

For Procrustes residuals and local shape variables the effects of isometric variation were removed by normalizing each individual to unit centroid size.
However, allometric relationships still influence covariance or correlation structure.
In order to remove this effect we used a procedure based upon @mitteroecker_comparison_2004, which relies on the estimation of an allometric component $a$ for each OTU, composed of normalized regression coefficents for each of the $m$ shape variables (either Procrustes residuals or local shape variables) over log Centroid Size.
If $\mathbf{S}$ is a covariance matrix, we obtained a covariance matrix $\mathbf{R}$ without the influence of allometric relationships using the equation
\begin{equation}
\mathbf{R} = (I_m - aa^t) \mathbf{S} (I_m - aa^t)
\label{eq:allo}
\end{equation}
where $I_m$ represents the identity matrix of size $m$.
Therefore, our empirical dataset consists of six sets of covariance/correlation matrices, corresponding to each type of morphometric variables considering the presence or absence of size variation.

## Empirical Tests

Using these six sets of covariance/correlation matrices, we tested the hypotheses of trait associations described in \autoref{tab:lms} for Procrustes residuals and \autoref{tab:dist} for interlandmark distances and local shape variables.
These trait sets are grouped with respect to their scope; two regional sets (Face and Neurocranium), each divided into three localized trait sets (Oral, Nasal and Zygomatic for the Face; Orbit, Base and Vault for the Neurocranium). 

For all hypotheses, we estimated Modularity Hypothesis Indexes [MHI; @porto_size_2013] and the RV coefficient [@klingenberg_morphometric_2009].
Both statistics are estimated by partitioning covariance or correlation matrices into blocks; if $\mathbf{A}$ is a covariance or correlation matrix, the partition
$$
\mathbf{A} =
\begin{bmatrix}
\mathbf{A}_h & \mathbf{A}_b \\
\mathbf{A}^t_b & \mathbf{A}_c
\end{bmatrix}
$$
indicates that the block $\mathbf{A}_h$ contains covariances or correlations between traits that belong to the trait set being considered, while $\mathbf{A}_c$ represents the complementary trait set; $\mathbf{A}_b$ represents the block of covariances or correlations between the two sets.
Thus, covariance ($\mathbf{S}$) or correlation ($\mathbf{C}$) matrices can be partitioned into a similar scheme. We estimated MHI values using the equation
\begin{equation}
MHI = \frac {\bar{\rho}_{+} - \bar{\rho}_{-}} {ICV}
\label{eq:mi}
\end{equation}
where $\bar{\rho}_{+}$ represents the mean correlation in $\mathbf{C}_h$, $\bar{\rho}_{-}$ represents the mean correlation in the remaining sets (both $\mathbf{C}_b$ and $\mathbf{C}_c$), and $ICV$ is the coefficient of variation of eigenvalues of the associated covariance matrix, which is a measurement of the overall integration between all traits [@shirai_skull_2010].
We estimated RV coefficients for each hypothesis using the relationship
\begin{equation}
RV = \frac{tr(\mathbf{S}_{b}\mathbf{S}^t_{b})}{\sqrt{tr(\mathbf{S}_h \mathbf{S}_h)tr(\mathbf{S}_c \mathbf{S}_c)}}
\label{eq:rv}
\end{equation}
where $tr$ represents the sum of diagonal elements in any given matrix ($tr \mathbf{A} = \sum_i a_{ii}$).

The partitioning scheme outlined above assumes that the complementary trait set does not represent an actual hypothesis; however, we may choose to consider that both sets ($\mathbf{A}_h$ and $\mathbf{A}_c$) represent two distinct hypothesis.
The estimation of RV coefficients remains the same; however, MHI values are estimated considering that $\bar{\rho}_{+}$ is the average correlation in both $\mathbf{C}_h$ and $\mathbf{C}_c$, while $\bar{\rho}_{-}$ represents the average correlation only in $\mathbf{C}_b$.
In the case of the distinction between Facial and Neurocranial traits, we estimated MHI values in this manner, reporting values for this estimate under the denomination 'Neuroface', following @marroig_comparison_2001, along independent MHI estimates for each region.
Furthermore, since both Face and Neurocranium are two disjoint trait sets when any morphometric variable type is considered, RV coefficient values for either set are equal; therefore, a single RV value is reported for both regions, for each variable type.

In order to test the hypothesis that a trait set represents a variational module, we used a randomization procedure generating 1000 random trait sets with the same number of traits as the original set, calculating MHI and RV values for each iteration.
For each trait set and covariance/correlation matrix, we used these values to construct distributions for both statistics representing the null hypothesis that a given trait set is a random arrangement without meaningful relationships; we then compare this null distribution to the real value obtained.
For MHIs we consider this null hypothesis rejected when the real value is higher than the upper bound for the distribution, considering the significance level established; for RV coefficients, the null hypothesis is rejected when the real RV value is lower than the lower bound for the distribution, also considering significance level.
For Procrustes residuals the randomization procedure maintains coordinates within the same landmark together in each randomly generated trait set, following @klingenberg_quantitative_2001.

While the procedure for estimating significance for MHIs is derived from Mantel's [-@mantel_detection_1967] approach [as outlined by @cheverud_methods_1989], we chose to generate null distributions for MHI directly, instead of estimating matrix correlation values for both real and randomized matrices.
Estimated $p$-values in both cases remain the same, and the additional step of calculating matrix correlations would produce an unnecessary difference between the estimation of signficance for MHI and RV.

## Estimation of Error Rates

```{r theomat, child = 'theomat.Rmd', eval = thesis}
```

```{r header_ms, child = 'header_ms.Rmd', eval = !thesis}
```

If samples were generated from a $\mathbf{C}_r$ matrix, this represents a situation of a true null hypothesis for each test, since the correlation matrix used to produce the sample was generated by a permutation of the hypothesis being tested.
Therefore, testing hypotheses over $\mathbf{C}_r$ matrices allows the estimation of Type I error rates, or the proportion of cases in which a true null hypothesis is rejected, given a significance level.
In an adequate test, we would expect that both quantities, significance level and Type I error rate, will be identical.

The opposite case, when we sampled $\mathbf{C}_s$ matrices, represents a situation in which we know that the null hypothesis of either test is false, since we are testing the hypothesis that the partitioning scheme used to generate that particular matrix actually represents two variational modules.
Thus, we estimated Type II error rates, or the probability that a false null hypothesis is not rejected, given a significance level; here, we represent Type II error using the power for each test, by simply calculating the complementar probability to Type II error rate.
In an adequate test, we expect that power would rapidly reach a plateau when significance level is still close to zero, and further increasing $P(\alpha)$ will not produce a great increase in power.

Our estimates of power for both statistics should also be controlled for effect size, since sampled correlations may generate a correlation structure that is not detected due to small differences among within-set and between-set correlations.
For each correlation matrix sampled, we estimate squared between-set correlations ($b^2$), in order to use it as an estimate of effect size that is not directly associated with either MHI and RV metrics.
We expect that power for either tests decreases with increasing $b^2$ values, as effect size would also decrease.

### Software

All analysis were performed under R `r paste (version $ major, version $ minor, sep = '.')` [@r_core_team_r:_2015].
Source code for all analyses can be found at <http://github.com/wgar84>.
Previous tests indicated no differences between our estimation of empirical RV coefficients, based upon our own code, and estimates provided by MorphoJ [@klingenberg_morphoj:_2011].
In order to obtain symmetrical landmarks configurations, we used code provided by Annat Haber, available at <http://life.bio.sunysb.edu/morph/soft-R.html>.

# Results

## Empirical Tests

Tests performed using MHI for localized trait sets (Oral, Nasal, Zygomatic, Orbit, Base, and Vault; \autoref{fig:Func}a) detected a consistent pattern among OTUs for interlandmark distances and local shape variables.
In the first set the Oral subregion was detected as a modular partition, and, with size removed, the Vault subregion was also detected; both Orbit and Base region were not detected in any of these tests.
With local shape variables, Oral, Vault Nasal, and Zygomatic sets were detected consistently across OTUs; the removal of allometric variation affected only the detection of the Vault in some groups.
Furthermore, the Base sub-region was detected only in 3 of 42 tests performed over local shape variables, pooling together size and size-free correlation patterns.

For Procrustes residuals, the pattern of detection among sub-regions and OTUs was more inconsistent; for instance, the Base sub-region was detected in several OTUs, which contrasts this type of morphometric variable with the other two types.
Observing the actual MHI values, \autoref{fig:Func}a also indicates that Procrustes residuals had a low variance of this metric within each OTU, while interlandmark distances and local shape variables have a consistent pattern of variation, with lower values for the Base and, for interlandmark distances, Orbit trait sets, while the Oral, Nasal and Vault regions displayed higher values consistently.

```{r Func, echo = FALSE, fig.cap = captions $ fig.Func, fig.width = 15, fig.height = 15, message = FALSE, fig.pos='h'}
plot_grid(modcomp.Plots $ MI.Func + ylab('') +
          ggtitle('Modularity Hypothesis Index') +
          theme(text = element_text(size = 12),
                strip.text = element_text(size = 9),
                axis.text = element_text(size = 14)),
          modcomp.Plots $ RV.Func + ylab('') +
          theme(text = element_text(size = 12),
                axis.text = element_text(size = 14),
                strip.text = element_text(size = 9)),
          nrow = 2, labels = c('a', 'b'), label_size = 36)
```

Tests performed using RV coefficients (\autoref{fig:Func}b) show a more irregular pattern for each variable type.
When interlandmark distances were considered, most tests detected the Vault sub-region with size variation retained, and the Base sub-region when size variation was removed.
For Procrustes residuals, few tests were able to reject the null hypotheses, detecting only a handful of valid modular partitions.
Tests performed on local shape variables displayed the opposite behavior: almost all partitions were detected, regardless of whether allometric variation had been retained or removed.
Moreover, RV values displayed a pattern of marked variation among OTUs, more so than between values within each OTU; notably, *Macaca fascicularis* and *Papio anubis* showed RV values much higher than those estimated on remaining species.
Such pattern can be observed both on interlandmark distances with size retained and in Procrustes residuals.

With respect to regional trait sets (Face and Neurocranium), tests performed using MHI (\autoref{fig:Dev}a) indicate a pattern consistent with those found on localized sets (\autoref{fig:Func}a).
Considering interlandmark distances, Facial traits were detected as a valid modular partition both with size variation retained or removed, while Neurocranial traits were detected as a valid partition only when size variation was removed.
This pattern mirrors the contrast between Oral and Vault traits found in the localized sets regarding interlandmark distances.
For Procrustes residuals, Neurocranial traits were detected as a valid partition with both size retained and removed; once again, this pattern mirrors the detection of the Basicranial partition found in the localized sets.
Finally, in local shape variables, both Face and Neurocranium were detected as valid with size retained; with size removed, only the Face was recognized consistently.
The same can be observed for localized sets, where the removal of allometric variation affected the detection of the Vault set in some OTUs.
The tests for the distinction of within-set and between-set correlations for these two sets (designated 'Neuroface') show a pattern that is consistent with tests for the individual sets: if one of the sets was previously detected, this distinction was also detected as valid.

```{r Dev, echo = FALSE, fig.cap = captions $ fig.Dev, fig.width = 15, fig.height = 10, message = FALSE, fig.pos='t'}

plot_grid(modcomp.Plots $ MI.Dev + ylab('') +
          ggtitle('Modularity Hypothesis Index') +
          theme(strip.text.y = element_text(size = 6),
                axis.text = element_text(size = 13)),
          modcomp.Plots $ RV.NeuroFace +
          theme(axis.text = element_text(size = 13)),
          nrow = 2, rel_heights = c(1.5, 1),
          labels = c('a', 'b'), label_size = 44)

```

Testing the distinction between Face and Neurocranium using RV coefficients (\autoref{fig:Dev}b) showed that in most cases both regions are considered distinct and valid variational modules for interlandmark distances and local shape variables; for Procrustes residuals, only a handful of taxa indicated the same result.
In this case, the correspondence with localized trait sets (\autoref{fig:Func}b) is more difficult due to the lack of independent tests for each region.

## Error Rates

Comparing the distributions of MHI and RV values from theoretical matrices with respect to their structure (\autoref{fig:stat_dist_sim}) indicated marked differences between metrics and morphometric variables from which correlations were sampled.
In general, the distribution of MHI values obtained from $\mathbf{C}_r$ matrices was the same regardless of type of variable, while the distributions for $\mathbf{C}_s$ matrices for this metric were heterogeneous; for either interlandmark distances and local shape variables, the distribution of MHI values for these two types of matrices did not overlap, while for Procrustes residuals, these distributions overlapped to some degree.
For RV values, all distributions overlap to some degree; for local shape variables and interlandmark distances with size retained, either distributions ($\mathbf{C}_r$ and $\mathbf{C}_s$) overlapped to a lesser extent.

```{r stat_dist_sim, echo = FALSE, fig.cap = captions $ fig.stat_dist_sim, fig.width = 6, fig.height = 6, message = FALSE, fig.pos='t'}
alt.dist.sim + theme (legend.text = element_text(size = 8),
                      axis.text = element_text(size = 8))
```

The relationship between significance levels and Type I error rates estimated over $\mathbf{C}_r$ matrices approached an identity relationship (\autoref{fig:type1}), regardless of whether variational modularity was estimated using MHI or RV; even at low sample sizes Type I error rates were very close to significance levels.
Furthermore, the effect of sampling correlations from size-free distributions did not change Type I error rates.

```{r type1, echo = FALSE, fig.cap = captions $ fig.type1, fig.width = 7, fig.height = 6, message = FALSE}
modsim.Plots $ Type1 + theme (legend.text = element_text(size = 8),
                              axis.text = element_text(size = 8),
                              strip.text = element_text(size = 8))
```

With respect to the relationship between power and significance levels (estimated over $\mathbf{C}_s$ matrices), there were substantial differences with respect to the chosen metric (MHI or RV) and to the type of variable that provides sampled correlations.
Considering local shape variables (\autoref{fig:type2_def}), tests using either MHI and RV implied high power, even at low sample or effect sizes; increasing these quantities further increased power.
However, for lower effect sizes (represented by high average squared correlation between sets, $b^2$) power for tests using MHI was higher than for those using RV; as effect size increased (lower $b^2$ values), the difference in power between the two statistics decreased. For local shape variables, sampling from its associated size-free correlation distribution implied minor differences in power for both statistics.

```{r type2_def, echo = FALSE, fig.cap = captions $ fig.type2_def, fig.width = 7, fig.height = 6, message = FALSE, fig.pos='t'}
modsim.Plots $ Type2.Def + theme (legend.text = element_text(size = 8),
                                  axis.text = element_text(size = 8),
                                  strip.text = element_text(size = 8)) +
  scale_color_brewer(name = 'Squared Correlation\nBetween Sets', palette = 'Set1')
```

For interlandmark distances (\autoref{fig:type2_ed}) there were substantial differences on the relationship between power and significance level.
In general, power for tests using MHI are always higher than for tests using RV; this effect is more pronounced on $\mathbf{C}_s$ matrices derived from size-free interlandmark distances, although tests on these matrices have a substantial decrease in power for either tests.
However, this decrease is more pronounced for tests based on the RV statistic, since for lower effect sizes (high $b^2$ values) power approaches an identity relationship with significance level.
Sample size also interferes with this relationship since increasing this quantity also increases power when higher effect sizes (low $b^2$ values) are considered.

```{r type2_ed, echo = FALSE, fig.cap = captions $ fig.type2_ed, fig.width = 7, fig.height = 6, message = FALSE, fig.pos='t'}
modsim.Plots $ Type2.ED + theme (legend.text = element_text(size = 8),
                                  axis.text = element_text(size = 8),
                                  strip.text = element_text(size = 8)) +
  scale_color_brewer(name = 'Squared Correlation\nBetween Sets', palette = 'Set1')
```

With respect to Procrustes residuals (\autoref{fig:type2_sym}), tests using either MHI or RV have reduced power regardless of effect or sample size. Sampling from size-free correlation distributions to build $\mathbf{C}_s$ matrices also has little effect.
In this case, power for tests performed using RV values approaches an identity relationship with significance level; increasing sample size has some effect, but it does not increase power above 50\% in any case.

```{r type2_sym, echo = FALSE, fig.cap = captions $ fig.type2_sym, fig.width = 7, fig.height = 6, message = FALSE, fig.pos='t'}
modsim.Plots $ Type2.Sym + theme (legend.text = element_text(size = 8),
                                  axis.text = element_text(size = 8),
                                  strip.text = element_text(size = 8)) +
  scale_color_brewer(name = 'Squared Correlation\nBetween Sets', palette = 'Set1')
```

# Discussion

Covariance matrices derived from morphological traits are supposed to represent the pattern of codependence that stems from interactions among developing morphological elements [@olson_morphological_1958; @cheverud_developmental_1996].
Such interactions are the expression of local developmental factors, as they interact with the gene expression profiles of surrounding cell types, producing coordinated changes in cellular growth and tissue interaction, thus integrating these elements in the adult population.
Although these events of local integration overlap, and the composed effect over adult covariance patterns may be confusing [@hallgrimsson_mouse_2008; @hallgrimsson_deciphering_2009], we believe that a careful comparison of different yet equally proper ways of measuring and representing form may be informative of the underlying processes that produce covariances.

Due to the minimization of quadratic distances among homologous landmarks during GPA, covariance matrices derived from Procrustes estimators lose the signal of localized effects on covariance patterns [@linde_inferring_2009].
Therefore, the use of Procrustes estimators to investigate morphological integration or modularity implicitly implies a dissociation between the phenomenom under investigation and the chosen representation.
This disconnection between theory and measurement may have serious consequences for the hypotheses we wish to test [@houle_measurement_2011]; such consequences are observable in both our empirical tests and those tests performed on theoretical matrices, as we explore below.

The mammalian Basicranium originates from thirteen precursor tissues derived from both paraxial mesoderm and neural crest, and they may merge to form single bones, such as the sphenoid [@jiang_tissue_2002; @lieberman_epigenetic_2011].
Furthermore, these precursors display a mosaic of endochondral and intramembranous ossification early in development, and, as the brain grows, it induces a pattern of internal resorption and exterior deposition on the underlying posterior Basicranium [@lieberman_primate_2000]; meanwhile, the anterior portion suffers influence from the development of Facial elements [@bastir_hierarchical_2005].
Thus, since the Basicranium ossifies early in development, the composed effect of all posterior steps of cranial development will overshadow any pattern of integration this region might have, as predicted by the palimpsest model of development [@hallgrimsson_deciphering_2009].
Moreover, the angulation between anterior and posterior elements of the Basicranium has significantly changed during primate evolution, and such property appears to have evolved in coordination with Facial growth relative to the cranial Vault, accomodating both structures on each other [@scott_cranial_1958; @lieberman_primate_2000; @lieberman_spatial_2008].

<!--- correlações do basicrânio são próximas de zero -->

Due to this heterogeneity of developmental processes acting on the Basicranium, it is not expected that this region would be detected as a variational module; thus, a test of this property over skull covariance patterns should fail to reject the null hypothesis of random association.
However, considering the 42 tests performed over covariance matrices derived from Procrustes residuals regarding localized hypotheses (midpanels of \autoref{fig:Func}a), the Basicranium is detected as a valid variational module in 27 cases, distributed through matrices with size either retained or removed; in some cases (e.g. *Alouatta*, *Cercopithecus*), only the Basicranium is detected.
Hence, Procrustes residuals show a pattern of detection of variational modules opposed to the expectation based on developmental dynamics.
In covariances matrices derived from interlandmark distances (upper panels of \autoref{fig:Func}a), the Basicranium was never detected; with local shape variables (lower panels of \autoref{fig:Func}a), the Basicranium was detected only three out of 42 times. 

These two remaining types of representations (i.e. interlandmark distances and local shape variables) show patterns of detection of variational modules that are consistent with the expectations derived from developmental and functional interactions, especially with respect to how these patterns articulate with the retention or removal of size variation.
Considering that interlandmark distances are on a ratio scale [@houle_measurement_2011], isometric variation will be represented to a greater extent when compared to subtle allometric relationships, due to the multiplicative nature of biological growth [@huxley_problems_1932].
Therefore, the Oral trait set is detected as a valid variational module, considering covariance matrices among interlandmark distances (upper left panel of \autoref{fig:Func}a), since this region is strongly affected by the induction of bone growth due to muscular activity beginning in the pre-weaning period [@zelditch_ontogenetic_1989; @hallgrimsson_deciphering_2009]. Furthermore, in these matrices, allometric interactions are associated with the strength of association between traits in Oral, Nasal, Zygomatic and Vault regions in an evolutionary scale (Chapter 3); thus, allometric relationships also play some role in the determination of covariance/correlation patterns for other subregions, although such pattern may be masked in interlandmark distances by the effect of isometric variation.
On the other hand, the patterns expressed in covariance/correlation matrices for local shape variables are influenced by allometric relationships defined on a proper scale [@jolicoeur_multivariate_1963; @houle_measurement_2011; @marquez_measurement_2012] and local developmental processes. Thus, they reflect this association between integration and allometry.

By removing allometric effects from local shape variables (lower right panel of \autoref{fig:Func}a), variational modularity can still be detected in both Oral and Nasal regions, while in a number of species, the Vault region is no longer detected as a variational module.
Vault integration may be achieved through both allometric relationships and the effect of relative brain growth, since Vault elements arise mostly through intramembranous ossification, induced by the secretion of signaling factors from the outer brain tissues, with a modest but necessary contribution of mesoderm-derived tissue that undergoes endochondral ossification [@jiang_tissue_2002; @rice_molecular_2003; @franz-odendaal_epigenetics_2011; @lieberman_epigenetic_2011].
In humans, the growth of Vault osteological elements occur, from a topological standpoint, without deviations from an isometric growth model, forming regular connections among bones [@esteve-altava_beyond_2014], since the boundary of interactions between brain and Vault bones is also regular.
Thus, as this effect dominates the latter stages of pre-natal development [@hallgrimsson_mouse_2008; @lieberman_epigenetic_2011], the overall effect of brain growth over skull growth patterns at this stage may mirror an affine transformation, rendering it undetectable over covariance patterns derived from local shape variables, since their estimation excludes shape variation associated with affine transformations [@marquez_measurement_2012].
In contrast, Oral and Nasal elements have more complex patterns of connectivity arising from their tight integration with the soft tissues that compose the remaining elements of the Face [@lieberman_epigenetic_2011; @esteve-altava_grist_2013; @esteve-altava_beyond_2014], thus producing an intrincate pattern of associations that may be responsible for the variational modularity we are able to detect in all species, regardless of whether size variation is retained or removed.

The differences between the pattern of detection in all representations for local trait sets for RV values (\autoref{fig:Func}b) makes a similar interpretation of the results for this metric difficult, as opposed to the results regarding MHI values.
The variance of RV values among species seems to indicate that RV values are sensitive to the magnitude of morphological integration, since species with high RV values, such as *Papio anubis* and *Macaca fascicularis*, are within genera that also exhibit higher magintudes among Catarrhines [@oliveira_covariance_2009], at least when we consider interlandmark distances.
Interestingly enough, this effect of magnitude of integration [which is thought to emerge as a consequence of size variation; @wagner_eigenvalue_1984; @marroig_comparison_2001; @marroig_size_2005; @porto_size_2013] seems to also affect Procrustes residuals regardless of whether covariance structure arising from allometric relationships is removed or not.

For the set of hypotheses concerning differences between Facial and Neurocranial traits (\autoref{fig:Dev}), there is substantial agreement between tests performed using MHI and RV values.
For MHI values (\autoref{fig:Dev}a), the overall pattern of detection is similar to the pattern detected in local trait sets, for all variable types; for RV values (\autoref{fig:Dev}b), there is strong support for the hypotheses that both Face and Neurocranium represent variational modules, in both local shape variables and interlandmark distances. 
These regions have marked differences in developmental timing and pattern formation, as observed from the behavior of their composing units [@zelditch_ontogenetic_1989; @hallgrimsson_deciphering_2009; @lieberman_epigenetic_2011; @esteve-altava_beyond_2014]; therefore, the more general pattern of distinction between Face and Neurocranium is detected regardless of the metric chosen to represent modularity.

## Theoretical Matrices and Error Rates

The distribution of MHI and RV values obtained from the theoretical matrices  (\autoref{fig:stat_dist_sim}) is a starting point for understanding the differences in power for tests using these two metrics (Figures \ref{fig:type2_def}--\ref{fig:type2_sym}). For MHI values, the distribution obtained from random ($\mathbf{C}_r$) matrices is consistently the same, regardless of what representation we used to sample correlations, or whether size was retained or removed.
On the other hand, the distribution of RV values for random matrices change depending on the representation sampled or whether size variation has been removed or retained.
Moreover, this change in behavior for the distribution of RV values for $\mathbf{C}_r$ matrices implies different levels of overlap between these null distributions and the distribution of values obtained for structured ($\mathbf{C}_s$) matrices.

For Procrustes residuals, a substantial overlap occurs regardless of whether size variation was removed or retained; unsurprisingly, our estimates of power for RV in this type of representation are very low (\autoref{fig:type2_sym}). Nonetheless, power estimated for tests based on MHI is lower than in other types of representation, since the difference between within-set and between-set correlations in Procrustes residuals (\autoref{fig:cor_dist}) is the lowest of all representations.
A substantial overlap in RV distributions for $\mathbf{C}_r$ and $\mathbf{C}_s$ matrices also occurs with interlandmark distances when size variation is removed, and it results in low power for tests using RV values in this type of representation (right column of \autoref{fig:type2_ed}).
However, in this case there is a substantial difference between within-set and between-set correlations (\autoref{fig:cor_dist}), and tests using MHI to represent modularity are still able to detect such difference (albeit with reduced power) when compared to tests over $\mathbf{C}_s$ matrices derived from interlandmark distances with size retained (\autoref{fig:type2_ed}).
In those cases where both distributions for $\mathbf{C}_r$ and $\mathbf{C}_s$ matrices do not overlap substantially --- for example, when local shape variables are considered (\autoref{fig:type2_def}) --- power for tests performed using MHI values is always higher than for tests using RV except when sample sizes are very high; in this case power for both metrics are similar.
The same behavior can also be observed in interlandmark distances when size is retained (left column of \autoref{fig:type2_ed}). 

These results indicate that RV coefficents are more sensitive to the absolute value of both within-set and between-set correlation distributions than MHI values.
For interlandmark distances (\autoref{fig:type2_ed}), removing size variation reduced the average value of both correlation distributions by a similar amount (\autoref{fig:cor_dist}); the difference between average correlations in these two sets actually increases, going from 0.042 to 0.061 when size is removed.
However, since the actual average correlations for these two distributions approach zero, tests based on RV lose power more rapidly than tests based on MHI. This sensitivity might be associated to the use of squared covariances, as shown by Equation \ref{eq:rv}, while Modularity Hypothesis Indexes use correlations directly (Equation \ref{eq:mi}).
Furthermore, as observed by @fruciano_resampling-based_2013, sample sizes influence the estimation of RV coefficients, and we demostrate here that such sensitivity also extends to estimates of power for tests using this metric.

Our estimates of power for tests using MHI indicate that it is more robust to differences in absolute correlation values or sample sizes, thus allowing comparisons across more heterogeneous settings, such as our comparison between different representations of form, with substantial variation in sample sizes.
The detection of variational modularity is akin to Student's *t* test, since we try to determine whether two groups of observations (correlations between traits in the same subset *versus* correlations between traits in different subsets) have a significant difference in average values; we use resampling procedures to estimate significance in this case only due to the interdependency between pairwise correlations.
Thus, as the estimation of MHI values resembles the estimation of *t*-values (difference between location parameters for two groups, divided by a scale parameter --- ICV; Equation \ref{eq:mi}), we believe that this statistic is appropriate to represent variational modularity, and that is reinforced by the robustness of the tests which employed MHI.

Our approach for constructing theoretical matrices attempts to simulate the simplest situation: only two subsets of traits, akin to the distinction between Face and Neurocranium in our empirical dataset (\autoref{fig:Dev}).
In this setting, both statistics are capable of detecting this distinction, except when both are used on covariance/correlation patterns derived from Procrustes residuals.
However, even though we built theoretical matrices using correlations sampled from these estimators, to simulate the interference in covariance structure that such estimators produce in our theoretical matrices is quite difficult.
Furthermore, constructing such matrices with more complicated patterns (with three modules, for instance) while maintaining their connection to the correlation distributions of each morphometric type is also difficult, due to the restriction on positive-definiteness we enforce on them.
Thus, the lack of differences in type I error rates for all cases may be a limitation of our scheme for building theoretical matrices.

The issues we found with the use of Procrustes estimators for covariance matrices and the use of RV coefficients to estimate and detect variational modularity may explain results found by other authors. For instance, @martinez-abadias_pervasive_2011 has found no evidence that genetic and phenotypic covariance structure for human skulls conforms to a modular structure, since all tests performed by these authors failed to reject the null hypothesis of random association.
These authors use Procrustes estimators to represent covariance structure, and test their hypothesis of partitioning (Face, Vault and Base) using RV as the statistic representing variational modularity.
Since this combination yields very low power (\autoref{fig:type2_sym}), not rejecting the null hypothesis in their case might be a consequence of the choice of estimates of both covariance structure and variational modularity; thus, these authors' conclusion of pervasive genetic integration in the human skull might be misleading, considering that skull covariance patterns in humans are one of the most modular examples of such patterns when compared to other mammals [@porto_evolution_2009] or catarrhine primates [@oliveira_covariance_2009].

The approach we explore in the present work is only one of the different alternatives to investigate the association between genetic, functional and developmental interactions and correlation structure [@mitteroecker_conceptual_2007].
For example, @perez_searching_2009 relies on abstracting correlation matrices into networks, then using community-detection algorithms to search for modular patterns without *a priori* hypotheses, associating their results with knockout experiments that support the communities they found among traits; however, it is not clear how much relevant biological information is retained in these network representations.
Furthermore, the authors use Procrustes estimators, which may bias the detection of modularity patterns in this setting in the same manner as we demonstrated here.

@monteiro_evolutionary_2005 assumes that the underlying morphogenetic components of the rodent mandible behave as modules, further investigating the patters of correlation between these units in both within-species variation and between-species variation among Echimids; @monteiro_adaptive_2010, relying on the correspondence of these units through mammalian diversification did the same to phylostomid bats.
Although using a landmark-based approach to represent morphological variation, the authors do not use Procrustes estimators to represent covariance structure among these units, and the pattern of reorganization of correlation structure among these units associated with niche diversification in phylostomids seems robust, considering that this radiation may have been associated with a very heterogeneous adaptive landscape, and such heterogeneity may lead to a reorganization of correlation patterns [@jones_mutation_2007; @jones_effects_2012; @melo_directional_2015].

Another valid approach is to model certain aspects of development as null hypotheses; @esteve-altava_beyond_2014, investigating the pattern of connections among human cranial bones, formulates the null hypothesis that unconstrained bone growth is sufficient to explain the observed patterns, further constructing a null network based on this expectation which is compared to the actual network.
Such approach could be extended to investigate covariance structure; considering the geometric properties of the features we measure [@mitteroecker_conceptual_2007], one could formulate the null hypothesis that topological proximity is a sufficient explanation for the observed covariance structure, against the alternative hypothesis that local developmental processes coupled with functional interactions produce stronger relationships among close elements that surpass topological interactions.
Alternatively, one could actively look for variational boundaries between regions, as boundary formation is a phenomenom that has been well studied under a dynamical perspective on development [e.g. @turing_chemical_1952; @meinhardt_boundary_1983; @tiedemann_dynamic_2012].

The approach of partitioning covariance matrices into blocks that correspond to inferred modular associations has the advantage that it is operationally simple; however, modularity patterns are almost certainly not expressed in phenotypic data as the binary hypotheses we used here [@hallgrimsson_deciphering_2009].
Thus, hypotheses and inferences derived from them have to be contextualized in the light of developmental dynamics, since the measurements we make and the parameters we estimate have to be properly connected to the models we are considering; otherwise, inferences made from such models may be devoid of meaning [@wagner_measurement_2010; @houle_measurement_2011].

## Conclusion

Here we show that Procrustes estimators for covariance matrices fail to capture the modularity patterns embedded in phenotypic data, regardless of which metric is chosen to represent such patterns, although the combination of this type of variable with RV coefficients for investigating modularity has even more problems than either has alone.
<!---
Considering the widespread use of this combination, at large a result of the compilation of analyses provided by the current geometric morphometric toolboxes available, our results become a cautionary tale about the dangers of using methods and analyses one does not fully grasp.
-->
Both interlandmark distances and local shape variables seem valid options to represent morphological variation, if their limitations are taken into consideration.
We wish to stress this point: any representation of morphological variation has limitations since they are themselves models --- at the very least of what it is important to represent --- not fully capturing the phenomena we may be interested in.


