---
output:
  pdf_document:
    includes:
      in_header: Tex/tese_head.tex
      before_body: Tex/tese_open.tex
      after_body: Tex/tese_app.tex
    fig_caption: true
    keep_tex: true
  word_document:
    fig_caption: true
    fig_width: 6
documentclass: report
classoption: twoside
fontsize: 11pt
csl: Bib/journal-of-evolutionary-biology.csl
bibliography: Bib/tese.bib
---	

```{r options, include = FALSE}
opts_chunk $ set (fig.path = 'Figures/', dev = 'pdf')
```

\pagestyle{fancy}

\renewcommand{\chaptermark}[1]{\markboth{#1}{}}
\renewcommand{\sectionmark}[1]{\markright{#1}{}}

\fancyhf{}
\fancyhead[RO]{\leftmark}
\fancyhead[LE]{\rightmark}
\fancyfoot[C]{\thepage}

\renewcommand{\headrulewidth}{0.5pt}

\newpage
\chapter{Introdução}
\label{ch:intro}
```{r chapter_intro, child = 'intro.Rmd'}
```

\def\sectionautorefname{Section}
\def\chapterautorefname{Chapter}
\def\figureautorefname{Figure}
\def\tableautorefname{Table}

\selectlanguage{english}

\newpage
\chapter{Error Rates in Modularity Hypothesis Testing: Comparing Linear and Geometric Morphometrics}
\label{ch:modcomp}
```{r chapter_modcomp, child = 'modcomp.Rmd'}
```

\newpage
\chapter{Allometric Constraints and Modularity}
\label{ch:allo}
```{r chapter_allo, child = 'allo.Rmd'}
```

\newpage
\chapter{A Phylogenetic Analysis of Shape Covariance Structure}
\label{ch:ppca}
```{r chapter_ppca, child = 'ppca.Rmd'}
```

\def\sectionautorefname{Seção}
\def\chapterautorefname{Capítulo}
\def\figureautorefname{Figura}
\def\tableautorefname{Tabela}

\selectlanguage{brazilian}

\pagestyle{plain}

\chapter*{Referências Bibliográficas}
\addcontentsline{toc}{chapter}{Referências Bibliográficas}
